---
title : Advent of Code 2022
subtitle: "[adventofcode.com](https://adventofcode.com/){.aoc-yellow}"
#author: Jo Dudding
#date: today
format: 
  html:
    #embed-resources: true
    #standalone: true
    fig-width: 8
    fig-height: 4
    theme: ['advent-of-code.scss']
    highlight-style: nord
    toc: true
    toc_float: 
      collapsed: true
    code-fold: true  
execute: 
  echo: true
  warning: false
  message: false
knitr: 
  opts_knit: 
    root-dir: "`r here::here()`"
---

# Introduction

This analysis is my attempt at [adventofcode.com](https://adventofcode.com/) using R and Quarto.

Here's an overview from [adventofcode.com](https://adventofcode.com/).

> Santa's reindeer typically eat regular reindeer food, but they need a lot of magical energy to deliver presents on Christmas. For that, their favorite snack is a special type of [star]{.aoc-yellow} fruit that only grows deep in the jungle. The Elves have brought you on their annual expedition to the grove where the fruit grows.
> 
> To supply enough magical energy, the expedition needs to retrieve a minimum of [fifty stars]{.aoc-yellow} by December 25th. Although the Elves assure you that the grove has plenty of fruit, you decide to grab any fruit you see along the way, just in case.
> 
> Collect stars by solving puzzles. Two puzzles will be made available on each day in the Advent calendar; the second puzzle is unlocked when you complete the first. Each puzzle grants [one star]{.aoc-yellow}. Good luck!


```{r}
#| label: setup

library(tidyverse)
library(scales)
library(here)

stars <- 0

```

# Day 1

[Instructions](https://adventofcode.com/2022/day/1)

```{r}
#| label: read-day1-part1
df1 <- tibble(cals = readLines('inputs/day1-part1.csv')) |>
  mutate(
    cals = as.numeric(cals),
    row = row_number(),
    elf = case_when(
      row == 1 ~ row,
      is.na(cals) ~ row
    )
  ) |>
  fill(elf) |>
  filter(!is.na(cals)) |>
  group_by(elf) |>
  summarise(cals = sum(cals)) |>
  ungroup() |>
  arrange(-cals)

stars <- stars + 1
```

The maximum calories carried by an elf is
[`r comma(max(df1$cals))`]{.aoc-green}. [*]{.aoc-yellow}

```{r}
#| label: read-day1-part2

top_3_cals <- df1 |>
  head(3) |>
  summarise(cals = sum(cals)) |>
  pull(cals)

stars <- stars + 1
```

The calories carried by the top three elves is
[`r comma(top_3_cals)`]{.aoc-green}. [*]{.aoc-yellow}

# Day 2

[Instructions](https://adventofcode.com/2022/day/2)

```{r}
#| label: read-day2-part1
df2 <- tibble(strat = readLines('inputs/day2-part1.csv')) |>
  separate(strat, into = c('them', 'us')) |>
  mutate(
    shape = case_when(
      us == 'X' ~ 1,
      us == 'Y' ~ 2,
      us == 'Z' ~ 3
    ),
    outcome = case_when(
      them == 'A' & us == 'X' ~ 3,
      them == 'A' & us == 'Y' ~ 6,
      them == 'A' & us == 'Z' ~ 0,
      them == 'B' & us == 'X' ~ 0,
      them == 'B' & us == 'Y' ~ 3,
      them == 'B' & us == 'Z' ~ 6,
      them == 'C' & us == 'X' ~ 6,
      them == 'C' & us == 'Y' ~ 0,
      them == 'C' & us == 'Z' ~ 3
    ),
    score = shape + outcome
  ) 

stars <- stars + 1

```

The total points for this strategy are
`r comma(sum(df2$score))`.  [*]{.aoc-yellow}

```{r}
#| label: read-day2-part2

df2_p2 <- df2 |>
  mutate(
    outcome = us,
    us = case_when(
      them == 'A' & outcome == 'X' ~ 'Z',
      them == 'A' & outcome == 'Y' ~ 'X',
      them == 'A' & outcome == 'Z' ~ 'Y',
      them == 'B' & outcome == 'X' ~ 'X',
      them == 'B' & outcome == 'Y' ~ 'Y',
      them == 'B' & outcome == 'Z' ~ 'Z',
      them == 'C' & outcome == 'X' ~ 'Y',
      them == 'C' & outcome == 'Y' ~ 'Z',
      them == 'C' & outcome == 'Z' ~ 'X'
    ),
    shape_points = case_when(
      us == 'X' ~ 1,
      us == 'Y' ~ 2,
      us == 'Z' ~ 3
    ),
    outcome_points = case_when(
      outcome =='X' ~ 0,
      outcome =='Y' ~ 3,
      outcome =='Z' ~ 6
    ),
    score = shape_points + outcome_points
  ) 

stars <- stars + 1
```

The total points for the updated strategy are
`r comma(sum(df2_p2$score))`.  [*]{.aoc-yellow}


# Day 3

[Instructions](https://adventofcode.com/2022/day/3)

```{r}
#| label: read-day3-part1

scoring3 <- tibble(
  contents_item = c(letters, LETTERS),
  priority = 1:52
)

df3 <- tibble(contents = readLines('inputs/day3-part1.csv')) |>
  mutate(
    rucksack = row_number(),
    len = str_length(contents),
    contents_item = str_split(contents, "")
  ) |>
  unnest(contents_item) |>
  group_by(rucksack) |>
  mutate(
    compartment = if_else(row_number() <= len / 2, 'a', 'b')
  ) 
  
df3_1 <- df3|>
  distinct(rucksack, contents_item, compartment) |>
  group_by(rucksack, contents_item) |>
  filter(n() > 1) |>
  ungroup() |>
  distinct(rucksack, contents_item) |>
  left_join(scoring3, by = 'contents_item')

stars <- stars + 1

```

The total priority for the rucksacks is
`r comma(sum(df3_1$priority))`.  [*]{.aoc-yellow}

```{r}
#| label: read-day3-part2

df3_2 <- df3 |>
  mutate(
    trio = (rucksack - 1) %/% 3
  ) |>
  distinct(rucksack, contents_item, trio) |>
  group_by(trio, contents_item) |>
  filter(n()  == 3) |>
  ungroup() |>
  distinct(contents_item, trio) |>
  left_join(scoring3, by = 'contents_item') 

stars <- stars + 1
```

The total priority for the badges is
`r comma(sum(df3_2$priority))`.  [*]{.aoc-yellow}

<!-- 
# Day 4

[Instructions](https://adventofcode.com/2022/day/4)

```{r}
#| label: read-day4-part1



stars <- stars + 0

```


```{r}
#| label: read-day4-part2



stars <- stars + 0
```


# Day 5

[Instructions](https://adventofcode.com/2022/day/5)

```{r}
#| label: read-day5-part1



stars <- stars + 0

```


```{r}
#| label: read-day5-part2



stars <- stars + 0
```


-->

# Total stars

The total number of [stars]{.aoc-yellow} earned is
[`r stars`]{.aoc-green}.

# Source code

The source code can be found at [github.com/JoDudding/advent-of-code](https://github.com/JoDudding/advent-of-code).

```{bash}
#| label: quarto-preview-instruction
#| echo: false

# quarto preview index.qmd
# quarto render index.qmd

```